Automatic differentiation via off-policy particle filters

### 2023-10-16 mop/mop.tex
Updates from Overleaf

\newcommand\Time{N}
\newcommand\ttime{n}
\newcommand\Np{J}
\newcommand\np{j}
\newcommand\altNp{q}
\newcommand\prob{\mathrm{Prob}}
\newcommand\code[1]{\texttt{#1}}

@@

\section{DOP-$\alpha$}

Here, we consider a doubly off policy particle filter, meaning that both \code{rprocess} and \code{rmeasure} are computed at $\phi$ and particles are reweighted to correspond to $\theta$.

%%%%  DOP(alpha) PFILTER PSEUDOCODE
\begin{algorithm}[t!]
  \caption{\textbf{Doubly off policy (DOP-$\alpha$) SMC}:
    \label{alg:mop-alpha}
  }
  Initialize filter particles:
  simulate ${X}_{0,j}^{F,\theta}\sim {f}_{{X}_{0}}\left(\mydot\giventh{\theta}\right)$ for $j$ in $\seq{1}{J}$\;
  Initialize relative weights: $w^{F,\theta}_{0,j}= 1$ for $j$ in $\seq{1}{J}$
  \;
  \For{$n$ in $\seq{1}{N}$}{
    Prediction weights with discounting: $w_{n,j}^{P,\theta} = \big(w_{n-1,j}^{F,\theta}\big)^\alpha$ for $j\ \mathrm{in}\ \seq{1}{J}$
    \nllabel{dop-alpha:discount}
    \;
    Simulate for prediction:
    ${X}_{n,j}^{P,\theta}\sim {f}_{{X}_{n}|{X}_{n-1}}\big(\mydot|{X}_{n-1,j}^{F};{\phi}\big)$ for $j\ \mathrm{in}\ \seq{1}{J}$ \nllabel{dop-alpha:step1}
    \;
    Adjust weights: $\displaystyle w_{n,j}^{P,\theta} = w_{n,j}^{P,\theta} \times
    \frac{{f}_{{X}_{n}|{X}_{n-1}}\big({X}_{n,j}^{P,\theta}|{X}_{n-1,j}^{F};{\theta}\big)}{{f}_{{X}_{n}|{X}_{n-1}}\big({X}_{n,j}^{P,\theta}|{X}_{n-1,j}^{F};{\phi}\big)}$
    for $j\ \mathrm{in}\ \seq{1}{J}$
    \nllabel{dop-alpha:dproc}
    \;
    Evaluate measurement density:
    $g^{\theta}_{n,j}={f}_{{Y}_{n}|{X}_{n}}(y_{n}^{*}|{X}_{n,j}^{P,\theta}\giventh{\theta})$ for $j$ in $\seq{1}{J}$\;
    Before-resampling conditional likelihood: $\displaystyle L_n^{B,\theta,\alpha} = \frac{\sum_{j=1}^Jg^\theta_{n,j} w^{P,\theta}_{n,j}}{\sum_{j=1}^J  w^{P,\theta}_{n,j}}$
    \;
    Conditional likelihood under $\phi$: 
    $L_n^{\phi} = \frac{1}{J}\sum_{m=1}^{J}g^{\phi}_{n,m}$
    \nllabel{dop-alpha:Lphi}
    \;
    Normalize weights:
    $\displaystyle \tilde{g}^{\phi}_{n,j}= \frac{g^{\phi}_{n,j}}{JL_n^{\phi}}$
    for $j\ \mathrm{in}\ \seq{1}{J}$
    \;
    Apply systematic resampling to select indices $k_{1:J}$ with $\prob\big(k_{j}=m\big) =\tilde{g}^{\phi}_{n,m}$ \nllabel{dop-alpha:systematic}\;
    Resample particles: ${X}_{n,j}^{F,\theta}={X}_{n,k_{j}}^{P,\theta}$
    \;
    Filter weights corrected for resampling:
    $\displaystyle w^{FC,\theta}_{n,j}= w^{P,\theta}_{n,j} \times \frac{ g^{\theta}_{n,j}}{ g^{\phi}_{n,j}}$ for $j\ \mathrm{in}\ \seq{1}{J}$ \nllabel{dop-alpha:weight:update}
    \;
    Resample filter weights:
    $w_{n,j}^{F,\theta}= {w}_{n,k_{j}}^{FC,\theta}$
    for $j$ in $\seq{1}{J}$ \nllabel{dop-alpha:step2}
    \;
    After-resampling conditional likelihood: $\displaystyle L_n^{A,\theta,\alpha} = L_n^\phi \, \frac{\sum_{j=1}^J w^{F,\theta}_{n,j}}{\sum_{j=1}^J  w^{P,\theta}_{n,j}}$
    \;
  }
\end{algorithm}

\begin{itemize}
\item In order to match the direct derivative for a discrete \code{rprocess} with $\alpha=0$, we would carry out the discounting in step~\ref{dop-alpha:discount} after the weight adjustment in step~\ref{dop-alpha:dproc}.
  However, if we want to make $\alpha=0$ correspond to a reasonable estimate with no weight memory (i.e., strong mixing) then we should construct the algorithms as stated.
\item ${X}_{n,j}^{P,\theta}$ in fact depends on $\phi$ not $\theta$ in this algorithm. Rename?
\end{itemize}

>>>>

\newcommand\Time{N}
\newcommand\ttime{n}
\newcommand\Np{J}
\newcommand\np{j}
\newcommand\altNp{q}
\newcommand\prob{\mathrm{Prob}}

@@
